---
title: "Regression, Growth Modeling, and Market Analysis"
author: "Ananya Iyer"
date: "2025-04-03"
output: 
  pdf_document:
    fig_width: 6
    fig_height: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r echo=FALSE, message=FALSE}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(boot)
```

# **Problem 1: Regression Warm Up**
```{r echo=FALSE, message=FALSE}
creatinine <- read.csv("creatinine.csv")
```

## **Part A**

```{r regression, echo=FALSE, message=FALSE}
model <- lm(creatclear ~ age, data = creatinine)

ggplot(creatinine, aes(x = age, y = creatclear)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE, color = "blue") +
  labs(x = "Age (years)", y = "Creatinine Clearance Rate (mL/minute)",
       title = "Relationship between Age and Creatinine Clearance Rate")
```

From our regression model, we can calculate the expected creatinine clearance rate for a 55-year-old:

```{r echo=FALSE, message=FALSE}
# Extract coefficients
intercept <- coef(model)[1]
slope <- coef(model)[2]

# Calculate expected clearance rate for age = 55
expected_rate_55 <- intercept + slope * 55
expected_rate_55
```
The expected creatinine clearance rate is calculated using the linear regression equation fit to the data:
Creatinine clearance rate = intercept + (slope × Age)

For a 55-year-old, substitute 55 into the equation.

The regression equation is:
Creatinine clearance rate = `r round(intercept, 2)` + (`r round(slope, 2)` × Age)

For a 55-year-old, the expected creatinine clearance rate is `r round(expected_rate_55, 2)` mL/minute.


## **Part B**
```{r echo=FALSE, message=FALSE}
# The slope coefficient represents the change in clearance rate per year of age
slope
```
The creatinine clearance rate changes by `r round(slope, 2)` mL/minute per year of age. This negative value indicates that as age increases by one year, the creatinine clearance rate decreases by approximately `r abs(round(slope, 2))` mL/minute.

## **Part C**
To determine whose creatinine clearance rate is healthier for their age, I calculated the expected clearance rate for each person and compared their actual rates to these expectations.

```{r echo=FALSE, message=FALSE}
# Expected rate for a 40-year-old
expected_40 <- intercept + slope * 40
# Expected rate for a 60-year-old
expected_60 <- intercept + slope * 60

# Actual rates
actual_40 <- 135
actual_60 <- 112

# Calculate residuals (actual - expected)
residual_40 <- actual_40 - expected_40
residual_60 <- actual_60 - expected_60

# Print results
data.frame(
  Age = c(40, 60),
  Actual_Rate = c(actual_40, actual_60),
  Expected_Rate = c(expected_40, expected_60),
  Residual = c(residual_40, residual_60)
)
```

The residual represents how much higher a person's clearance rate is than expected for their age. The `r ifelse(residual_40 > residual_60, "40", "60")`-year-old has a residual of `r round(max(residual_40, residual_60), 2)` mL/minute, which is higher than the `r ifelse(residual_40 > residual_60, "60", "40")`-year-old's residual of `r round(min(residual_40, residual_60), 2)` mL/minute.

Therefore, the `r ifelse(residual_40 > residual_60, "40", "60")`-year-old has a healthier (higher) creatinine clearance rate for their age.

# **Problem 2: Modeling Disease Growth**
```{r echo=FALSE, message=FALSE}
covid <- read.csv("covid.csv")
```

## **Parts 1 and 2**

```{r fit_models, echo=FALSE, message=FALSE}
# Split data by country
italy_data <- subset(covid, country == "Italy")
spain_data <- subset(covid, country == "Spain")

#Fit exponential growth models for each country and calculate growth rates and doubling times.
# For Italy
italy_model <- lm(log(deaths) ~ days_since_first_death, data = italy_data)
italy_growth_rate <- coef(italy_model)[2]
italy_doubling_time <- log(2) / italy_growth_rate

# For Spain
spain_model <- lm(log(deaths) ~ days_since_first_death, data = spain_data)
spain_growth_rate <- coef(spain_model)[2]
spain_doubling_time <- log(2) / spain_growth_rate

# Summary
cat("Italy Growth Rate:", round(italy_growth_rate, 3), "\n")
cat("Italy Doubling Time:", round(italy_doubling_time, 1), "days\n")
cat("Spain Growth Rate:", round(spain_growth_rate, 3), "\n")
cat("Spain Doubling Time:", round(spain_doubling_time, 1), "days\n")
```


```{r bootstrap, echo=FALSE, message=FALSE}
#bootstrapping to get 95% confidence intervals
set.seed(123)
n_boot <- 1000

# Bootstrap function
bootstrap_exp_growth <- function(data) {
  n <- nrow(data)
  indices <- sample(1:n, n, replace = TRUE)
  boot_data <- data[indices, ]
  model <- lm(log(deaths) ~ days_since_first_death, data = boot_data)
  growth_rate <- coef(model)[2]
  doubling_time <- log(2) / growth_rate
  return(c(growth_rate, doubling_time))
}

# Bootstrap for Italy
italy_boot <- replicate(n_boot, bootstrap_exp_growth(italy_data))
italy_growth_ci <- quantile(italy_boot[1,], c(0.025, 0.975))
italy_doubling_ci <- quantile(italy_boot[2,], c(0.025, 0.975))

# Bootstrap for Spain
spain_boot <- replicate(n_boot, bootstrap_exp_growth(spain_data))
spain_growth_ci <- quantile(spain_boot[1,], c(0.025, 0.975))
spain_doubling_ci <- quantile(spain_boot[2,], c(0.025, 0.975))

# Results
cat("Italy Growth Rate:", round(italy_growth_rate, 3), 
    "95% CI: [", round(italy_growth_ci[1], 3), ",", round(italy_growth_ci[2], 3), "]\n")
cat("Italy Doubling Time:", round(italy_doubling_time, 1), "days",
    "95% CI: [", round(italy_doubling_ci[1], 1), ",", round(italy_doubling_ci[2], 1), "]\n")
cat("Spain Growth Rate:", round(spain_growth_rate, 3),
    "95% CI: [", round(spain_growth_ci[1], 3), ",", round(spain_growth_ci[2], 3), "]\n")
cat("Spain Doubling Time:", round(spain_doubling_time, 1), "days",
    "95% CI: [", round(spain_doubling_ci[1], 1), ",", round(spain_doubling_ci[2], 1), "]\n")
```

## **Part 3**

```{r line graph, echo=FALSE, message=FALSE}
#line graph showing reported daily deaths over time for each country
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line() +
  geom_point(alpha = 0.5) +
  scale_color_manual(values = c("Italy" = "darkred", "Spain" = "darkblue")) +
  labs(x = "Days Since First Death", y = "Daily Deaths",
       title = "COVID-19 Daily Deaths: Italy vs Spain (2020)") +
  theme_minimal() +
  theme(legend.title = element_blank())
```


# **Problem 3: Price Elasticity of Demand**

```{r, echo=FALSE, message=FALSE}
milk <- read.csv("milk.csv")
```

The power-law model is $Q = KP^{\beta}$, where $P$ is price, $Q$ is quantity, and $\beta$ is the price elasticity of demand.

Taking logarithms of both sides:
$\log(Q) = \log(K) + \beta \times \log(P)$

Fit this model using linear regression.

```{r elasticity, echo=FALSE, message=FALSE}
# Fit log-log model
log_model <- lm(log(sales) ~ log(price), data = milk)

# The coefficient for log(price) is the elasticity
elasticity <- coef(log_model)[2]
```

Bootstrap to get a confidence interval for the elasticity.

```{r, echo=FALSE, message=FALSE}
set.seed(456)
n_boot <- 1000

# Bootstrap function
bootstrap_elasticity <- function(data) {
  n <- nrow(data)
  indices <- sample(1:n, n, replace = TRUE)
  boot_data <- data[indices, ]
  model <- lm(log(sales) ~ log(price), data = boot_data)
  return(coef(model)[2])
}

# Bootstrap
elasticity_boot <- replicate(n_boot, bootstrap_elasticity(milk))
elasticity_ci <- quantile(elasticity_boot, c(0.025, 0.975))

# Results
cat("Price Elasticity of Demand:", round(elasticity, 3),
    "95% CI: [", round(elasticity_ci[1], 2), ",", round(elasticity_ci[2], 2), "]\n")
```

```{r plot_elasticity, fig.width=8, fig.height=5, echo=FALSE, message=FALSE}
# Visualize the relationship
ggplot(milk, aes(x = price, y = sales)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x, se = FALSE, color = "red") +
  scale_x_log10() +
  scale_y_log10() +
  labs(x = "Price (log scale)", y = "Quantity Sold (log scale)",
       title = "Log-Log Relationship Between Milk Price and Sales") +
  theme_minimal()
```

The estimated price elasticity of demand for milk is `r round(elasticity, 3)` with a 95% confidence interval of [`r round(elasticity_ci[1], 2)`, `r round(elasticity_ci[2], 2)`]. This value represents the percentage change in quantity demanded in response to a 1% change in price. To estimate this elasticity, I fit a log-log regression model where the logarithm of sales was regressed on the logarithm of price. The coefficient on log(price) gives us the elasticity estimate.

